---
title: "Calculating zeros and probabilities"
date: "`r Sys.Date()`"
output:  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calculating zeros and probabilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction 

The aim of this document is to illustrate how to calculate zeros and probabilities using the icesRDBES package.

```{r}
library(icesRDBES)
```


## Prerequisites

First we'll load some example data from the RDBES and check it's valid.  It's a good tip to check your RDBESRawObjects are valid after any manipulations you perform.

```{r load}

getwd()

# H7 test data
load("../tests/testthat/h7_v_1_19/testDataPreparationFunctions.Rdata")
myRawObject <- myH7RawObject

# Number of rows in each non-null table
unlist(lapply(myRawObject, nrow))

validateRDBESRawObject(myRawObject)


```


## Check selection and probability values

First we remove the prefixes and then check the selection and probability values for any issues - there are some things we need to fix.

```{r checkSelAndProbs, error = TRUE}


# checks data for some issues related to selection and probabilities
# note: this function is only exemplified here - 
# in general only applyGenerateProbs will be used as the wrapper also includes
# this function
runChecksOnSelectionAndProbs(myRawObject, printStopIssue = FALSE)



```

First, we'll change some data related to selection methods and stratification.

```{r changeData1}

	# changes to selection methods
	myRawObject$OS$OSselectMeth<-"SRSWOR"
	
	myRawObject$SA$SAstratification<-"N"
	
	myRawObject$SA$SAselectMeth<-"SRSWOR"
	
	myRawObject$BV$BVselectMeth<-"SRSWOR"
	
	runChecksOnSelectionAndProbs(myRawObject)

```
Now the main issues related to selection methods have been corrected (and documented) but still probabilities cannot be calculated because numTotal are missing. 


```{r changeData2, error = TRUE}

  myH7PrepObject<-applyGenerateProbs (x = myRawObject
                                      , probType = "inclusion"
                                      , overwrite=T
                                      , runInitialProbChecks = FALSE)
	

```
 
 We'll fix that now.
 
```{r changeData3, error = TRUE}

	# changes to numTotal
  myRawObject$OS$OSnumTotal<-c(7,7,7)
  myRawObject$SS$SSnumTotal<-c(1,1,1)
  myRawObject$SA$SAnumTotal<-c(25, 17, 32)
  myRawObject$BV$BVnumTotal<-c(rep(50, 29*5), rep(200, 95*5), rep(200, 39*5))
	
	myH7PrepObject<-applyGenerateProbs (x = myRawObject
	                                    , probType = "inclusion"
	                                    , overwrite=T
	                                    , runInitialProbChecks = FALSE)

```

We're still missing some numSamp values - let's add these and try again.


```{r changeData4}

  myRawObject$OS$OSnumSamp<-c(3,3,3)
  myRawObject$SS$SSnumSamp<-c(1,1,1)
  myRawObject$SA$SAnumSamp<-c(1, 1, 1)
  myRawObject$BV$BVnumSamp<-c(rep(50, 29*5), rep(200, 95*5), rep(200, 39*5))

  myH7PrepObject<-applyGenerateProbs (x = myRawObject
                                      , probType = "inclusion"
                                      , overwrite=T
                                      , runInitialProbChecks = FALSE)
	

```

Success! We now have a prepared object - let's check our object is still valid.

```{r checkValid1}
  validateRDBESRawObject(myH7PrepObject)
```


## Generate zeroes using SL

Generate true zeroes using the data from the SL table.

```{r generateZeroes}

  myH7PrepObjectWithZeroes<-generateZerosUsingSL(myH7PrepObject)

  myH7PrepObjectWithZeroes$SA

  # note the solution now has floating points values for SAid, and SAseqNum
  # (we are using sprintf to ensure the decimal places are displayed)
	sprintf(myH7PrepObjectWithZeroes[['SA']]$SAid, fmt = '%.3f')
	sprintf(myH7PrepObjectWithZeroes[['SA']]$SAseqNum, fmt = '%.3f')


```

Finally, let's check our object is still valid.

```{r checkValid2}
  validateRDBESRawObject(myH7PrepObjectWithZeroes)
```



```{r}
#END
```
