---
title: "Calculating zeros and probabilities"
date: "`r Sys.Date()`"
output:  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calculating zeros and probabilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction 

The aim of this document is to illustrate how to calculate zeros and probabilities using the icesRDBES package.

```{r}
library(icesRDBES)
```
## Load functions from their temporary locations

These should be added to the icesRDBES package which will make this step unnecessary in the future.

```{r loadFunctionsTemp}

# clean the table prefixes of variable names 
#[makes it easier to handle in other functions]
	source("R/removePrefixFromVarNames.R") 
# checks data for some issues related to selection and probabilities, 
# including some features not yet developed 
# [stops need to corrected before probabilities are calculated]
	source("R/runChecksOnSelectionAndProbs.R") 
# modified/improved version of generateProbs developed during WKRDB-EST 1 
# and WKRDB-EST 2 [several new functionalities, warnings, etc]
	source("R/generateProbs.r") 
 # wrapper to "runChecksOnSelectionAndProbs" and "generateProbs"
	source("R/applyGenerateProbs.r")
  source("R/generateZerosUsingSL.r")


```


## Prerequisites

First we'll load some example data from the RDBES and check it's valid.  It's a good tip to check your RDBESRawObjects are valid after any manipulations you perform.

```{r load}

getwd()

# H7 test data
load ("data/testDataPreparationFunctions.Rdata")

# Number of rows in each non-null table
unlist(lapply(myH7RawObject, nrow))

validateRDBESRawObject(myH7RawObject, verbose = TRUE)


```


## Check selection and probability values

First we remove the prefixes and then check the selection and probability values for any issues - there are some things we need to fix.

```{r checkSelAndProbs, error = TRUE}

# clean the prefixes [makes it easier to handle in other functions]
myH7RawObject <- removePrefixFromVarNames(myH7RawObject)

class(myH7RawObject)

# checks data for some issues related to selection and probabilities
# note: this function is only exemplified here - 
# in general only applyGenerateProbs will be used as the wrapper also includes
# this function
runChecksOnSelectionAndProbs(myH7RawObject, printStopIssue = FALSE)



```

First, we'll change the selection methods.

```{r changeData1}

	# changes to selection methods
	myH7RawObject$OS$selectMeth<-"SRSWOR"
	
	myH7RawObject$SA$stratification<-"N"
	
	myH7RawObject$SA$selectMeth<-"SRSWOR"
	
	myH7RawObject$BV$selectMeth<-"SRSWOR"
	
	runChecksOnSelectionAndProbs(myH7RawObject)

```
Now the main issues relation to selection methods have been corrected (and documented) but still probabilities cannot be calculated because numTotal are missing. 


```{r changeData2, error = TRUE}

  myH7PrepObject<-applyGenerateProbs (x = myH7RawObject
                                      , probType = "both"
                                      , overwrite=T
                                      , runInitialProbChecks = FALSE)
	

```
 
 We'll fix that now.
 
```{r changeData3}

	# changes to numTotal
	myH7RawObject$OS$numTotal<-c(3,3,3)
	myH7RawObject$SS$numTotal<-c(1,1,1)
	myH7RawObject$SA$numTotal<-c(20, 10, 5)
	myH7RawObject$BV$numTotal<-c(rep(50, 29*5), rep(200, 95*5), rep(200, 39*5))
	
	myH7PrepObject<-applyGenerateProbs (x = myH7RawObject
	                                    , probType = "both"
	                                    , overwrite=T
	                                    , runInitialProbChecks = FALSE)

```
## Generate zeroes using SL

Generate true zeroes using the data from the SL table.

```{r generateZeroes}

  myH7PrepObject2<-generateZerosUsingSL(myH7PrepObject)

# note the solution [dec SAid added to avoid integer that could 
# duplicate other already existing]
	myH7PrepObject2$SA


```

Finally, let's check our object is still valid. (It won't be until Nuno changes his function code :-) )

```{r checkValid}
  validateRDBESRawObject(myH7PrepObject2, verbose = TRUE)
```



```{r}
#END
```
