---
title: "Estimating Population parameters from RDBESDataObjects"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimating Population parameters from RDBESDataObjects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction 

The aim of this document is to illustrate how population parameters like total and mean can be estimated using the **estimMC(...)** function.

```{r}
library(RDBEScore)
```

## Prerequisites

First we'll load some example data from the RDBES and check it's valid.  It's a good idea to check your RDBESDataObjects are valid after any manipulations you perform.

```{r load}
# h1d <- importRDBESDataCSV(rdbesExtractPath = "../tests/testthat/h1_v_1_19_13")
# validateRDBESDataObject(h1d, verbose = FALSE)

```


## Single level Multiple Count Estimator

Let's first estimate the last level values for a single VSid so that estimation is done for each top level record (i.e. as the data is)
```{r}
# FMidSel <- "1445683"
# BV <- h1d$BV[h1d$BV$FMid == FMidSel,]
# 
# estimMC(as.numeric(BV$BVvalueMeas),
#         BV$BVnumSamp,
#         BV$BVnumTotal, 
#         method=unique(BV$BVselectMeth))
```




```{r}
#END
```

