---
title: "Raw Data Import"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Raw Data Import}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction 

The aim of this document is to outline the basic workflow of importing data downloaded from the [ICES Regional Database & Estimation System (RDBES)](https://sboxrdbes.ices.dk/#/) into `R` using the `RDBEScore` package.


## Load the package

```{r setup}
library(RDBEScore)
```

## Importing zipped files

```{r}
#directory holding the files
ddir <- "../tests/testthat/h1_v_1_19_13/"

#the files to import
zipFiles <- c("H1_2022_10_05.zip",
              "HSL_2022_10_05.zip",
              "HVD_2022_10_05.zip",
              "VesselSelection.csv")

#adding full path
zipFiles <- paste0(ddir, zipFiles)
```


It can directly import the `.zip` archive from the RDBES download containing all mandatory Hierarchy tables plus VD and SL:

```{r}
importedH1 <- importRDBESDownloadData("../tests/testthat/h1_v_1_19_13/H1_2022_10_05.zip")
#print the not NULL table names
names(importedH1[!unlist(lapply(importedH1, is.null))])
```

It can import the VD or SL tables `.zip` archives, but will include all other tables as `NULL`:

```{r}
importedSL <- importRDBESDownloadData("../tests/testthat/h1_v_1_19_13/HSL_2022_10_05.zip")
#print the not NULL table names
str(importedSL)
```

or a combination of both

```{r}
importedTables <- importRDBESDownloadData(zipFiles[1:3])
#print the not NULL table names
names(importedTables[!unlist(lapply(importedTables, is.null))])
```

It can also handle overwriting zip file original files with files apearing later in the list. However each overwrite results in a warning!


```{r}
importedTables <- importRDBESDownloadData(zipFiles)
#print the not NULL table names
names(importedTables[!unlist(lapply(importedTables, is.null))])
```


It can also import the unzipped `.csv` files with the default RDBES names:

```{r}
importedH5 <- importRDBESDataCSV("../tests/testthat/h5_v_1_19_13")
#print the not NULL table names
names(importedTables[!unlist(lapply(importedTables, is.null))])
```


## Object class

It should be noted that the objects created are of the S3 class "RDBESDataObject"

```{r}
importedTables <- importRDBESDownloadData(zipFiles)
class(importedTables)
```


RDBESDataObject structure can be validated using the validateRDBESDataObject() function.

```{r}
importedTables <- importRDBESDownloadData(zipFiles)
validateRDBESDataObject(importedTables, verbose = FALSE)
```


```{r}
#END
```


